---
title: "Machine Learning Assignment"
author: "Rajib Kar"
date: "May 1, 2017"
output: html_document
---
Background
----------

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

Data
-----

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment. 

Preprocessing and Loading the data
--------------------------------
After downloading the data, we have loaded the same in R .

we have removed zero divisions(Div/0) by NAs.

```{r,cache=TRUE}
training <- read.csv("pml-training.csv", na.strings=c("#DIV/0!"), row.names = 1)

testing <- read.csv("pml-testing.csv", na.strings=c("#DIV/0!"), row.names = 1)

```
Checking data
-------------
First 7 variables are for auditing hence removing them from training dataset.
```{r}
training<-training[,-c(1:7)]
```
Removal of near zero variables
------------------------------
It is necessary to remove the near zero or parse variables, it will reduce the number of predictors.

```{r}
###near zero value removal#######

thr<-dim(training)[1]*0.95

usfulcol<-!apply(training,2,function(x) sum(is.na(x))>thr|| sum(x=="")>thr)

training<-training[,usfulcol]

library(caret)

badCols <- nearZeroVar(training, saveMetrics = TRUE)

training <- training[, badCols$nzv==FALSE]

```
Preparing Training and Validation set
------------------------------------
Preparing the training and the validation datasets. We will use caret package for preprocessing.
```{r}
set.seed(1234)

inTrain<-createDataPartition(y=training$classe,p=.7,list = FALSE)

training<-training[inTrain,]

val<-training[-inTrain,]
```
Also checking the frequencies of different values of Variable Classe, which is required to be predicted.
```{r}
plot(training$classe, col="blue", main="Data frequency per activity"
     , xlab="classe value", ylab="Frequency")
```

We can see Classe A is more frequent than others.

Model Selection and Validation
--------------------------------
We will use rpart and Random forest and chose the best out of those two.

rpart
-----

```{r,cache=TRUE}
###Rpart###

library(rpart)
training$classe<-as.factor(training$classe)

mdrpart<-rpart(classe~.,data=training,method="class")

confusionMatrix(predict(mdrpart,val,type="class"),val$classe)
```

Random Forest
-------------

```{r}
library(randomForest)

mdrf<-randomForest(classe~.,data=training,importance=TRUE,ntrees=10)

confusionMatrix(predict(mdrf,val,type="class"),val$classe)
```
Conclusion
----------
It is clear from the outcome of the Confusion Matrix that the accuracy of Random forest is much greater than that of Rpart.So we will considering Random Forest over Rpart while predicting the outcome of test set.

Predicting Test set
--------------------
Now let us predict the test set.
```{r}
library(caret)

testpredict<-predict(mdrf,testing,type="class")

testpredict
```